---
title: "looming"
author: "Chloe Brown"
date: "04/10/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I have three sets of data here. The first is looming which compares an expanding black circle on white background with an expanding white circle on grey background. The second compares an expanding black circle on white background (positive control) with two stimuli with an even intensity as the two circles expand. The third set of data compiles both experiments into one. The two experiments were done several months apart.


```{r}
library(sciplot)
library(multcomp)
library(ade4)
library(vegan)
library(ggplot2)
library('tidyverse')
library(lme4)
library(ggeffects)
```

```{r}
loomingcomb <- read_csv("looming_comb.csv")
loomingcomb %>% filter(experiment=="first") %>% droplevels -> looming
loomingcomb %>% filter(experiment=="even") %>% droplevels -> loomingeven
```


#for first experiment with just black and white expanding circles
black = positive control
white = negative control
```{r}
looming$stimulus <- as.factor(looming$stimulus)
looming$chiton <- as.factor(looming$chiton)
looming$experiment <-as.factor(looming$experiment)
looming$response <- as.numeric(looming$response)


str(looming)
```

```{r}
#barplot
mean.countlooming <- tapply(looming$response, list(looming$stimulus), mean)
mean.countlooming
sd.count <- tapply(looming$response, list(looming$stimulus), sd)
sd.count
n.count <- tapply(looming$response, list(looming$stimulus), length)
n.count
se.count <- sd.count/sqrt(n.count)
se.count
mids <- barplot(mean.countlooming, ylim = c(0,1), xlab = "Stimulus", ylab = "Shadow Responses Elicited (%)") 
mids <- barplot(mean.countlooming, ylim = c(0,1), xlab = "Stimulus", ylab = "Shadow Responses Elicited (%)") +arrows(mids, mean.countlooming-se.count, mids, mean.countlooming+se.count,code = 3, angle = 90, length = 0.1)

```

#for second experiment with even intensity stimuli and black circle
```{r}
loomingeven$stimulus <- as.factor(loomingeven$stimulus)
loomingeven$chiton <- as.factor(loomingeven$chiton)
loomingeven$experiment <- as.factor(loomingeven$experiment)
loomingeven$response <-as.numeric(loomingeven$response)

str(loomingeven)
```
Even black is a black circle on white background that expands to a lighter one on a white background = ensure responses are to circle changing.
Even background is a white circle on a grey background expanding to a grey circle on a white background = check responses aren't due to overall changes in projection instead of just circle.
black = positive control
```{r}
#barplot
mean.counte <- tapply(loomingeven$response, list(loomingeven$stimulus), mean)
mean.counte
sd.counte <- tapply(loomingeven$response, list(loomingeven$stimulus), sd)
sd.counte
n.counte <- tapply(loomingeven$response, list(loomingeven$stimulus), length)
n.counte
se.counte <- sd.counte/sqrt(n.counte)
se.counte
midz <- barplot(mean.counte, ylim = c(0,1), xlab = "Stimulus", ylab = "Shadow Responses Elicited (%)") 
midz <- barplot(mean.counte, ylim = c(0,1), xlab = "Stimulus", ylab = "Shadow Responses Elicited (%)") +arrows(midz, mean.counte-se.counte, midz, mean.counte+se.counte,code = 3, angle = 90, length = 0.1)

```
<<<<<<< HEAD


#Combination of looming even and looming
Even black is a black circle on white background that expands to a lighter one on a white background = ensure responses are to circle changing.
Even background is a white circle on a grey background expanding to a grey circle on a white background = check responses aren't due to overall changes in projection instead of just circle.
black = positive control
white = negative control
=======
#Comparison of positive controls across two looming experiments.

```{r}
poscontrol <- subset(loomingcomb, stimulus == "Black")
poscontrol

bargraph.CI(experiment, response, legend=T, data=poscontrol, ylim=c(0,1), xlab = "Stimulus", ylab = "Response") #tp see black responses separately.

pos.model <- glmer(response ~ experiment + (1|chiton), data = poscontrol, family = binomial(),                             control = glmerControl(optimizer = "optimx", calc.derivs = FALSE,
                                              optCtrl = list(method = "nlminb", starttests = FALSE)))
pos.model.null <- glmer(response ~ 1 + (1|chiton), data = poscontrol, family = binomial(),                             control = glmerControl(optimizer = "optimx", calc.derivs = FALSE,
                                              optCtrl = list(method = "nlminb", starttests = FALSE)))

anova(pos.model,pos.model.null)
summary(glht(pos.model, mcp(experiment="Tukey")))
```
There's no SD between null and model and AIC is only very slightly better. Therefore, two experiments will be combined into one for analysis

```{r}
bargraph.CI(stimulus, response, legend=T, data=loomingcomb, ylim=c(0,1), xlab = "Stimulus", ylab = "Response") #tp see black responses separately.
```



```{r}
loomingcomb$stimulus <- as.factor(loomingcomb$stimulus)
loomingcomb$chiton <- as.factor(loomingcomb$chiton)
loomingcomb$experiment <-as.factor(loomingcomb$experiment)
loomingcomb$response <-as.numeric(loomingcomb$response)

str(loomingcomb)

mean.count <- tapply(loomingcomb$response, list(loomingcomb$stimulus), mean)
mean.count
sd.count <- tapply(loomingcomb$response, list(loomingcomb$stimulus), sd)
sd.count
n.count <- tapply(loomingcomb$response, list(loomingcomb$stimulus), length)
n.count
se.count <- sd.count/sqrt(n.count)
se.count

```


```{r}
bargraph.CI(stimulus, response, group = experiment, legend=T, data=loomingcomb, ylim=c(0,1), xlab = "Stimulus", ylab = "Response") #tp see black responses separately.

```

##Model

###Null Model
 
```{r}
loomingmodel.null <-glmer(response ~ 1 + (1|chiton), data = loomingcomb, family = binomial(),                             control = glmerControl(optimizer = "optimx", calc.derivs = FALSE,
                                              optCtrl = list(method = "nlminb", starttests = FALSE)))
```
 
#to see if there's an effect of length.
```{r}
loomingmodel.example <- glmer(response ~ stimulus * length + (1|chiton),
                              data = loomingcomb, family = binomial(),      
                    control = glmerControl(
                      optimizer = "optimx", calc.derivs = FALSE,
                                           optCtrl = list(method = "nlminb", starttests = FALSE)))  

loomingmodel.example2 <- glmer(response ~ stimulus + length + (1|chiton),
                              data = loomingcomb, family = binomial(),      
                    control = glmerControl(
                      optimizer = "optimx", calc.derivs = FALSE,
                                           optCtrl = list(method = "nlminb", starttests = FALSE)))  

loomingmodel.example3 <- glmer(response ~ stimulus + (1|chiton),
                              data = loomingcomb, family = binomial(),      
                    control = glmerControl(
                      optimizer = "optimx", calc.derivs = FALSE,
                                           optCtrl = list(method = "nlminb", starttests = FALSE)))  

anova(loomingmodel.example,loomingmodel.example2,loomingmodel.example3,loomingmodel.null)

summary(loomingmodel.example)

```
 

AIC comparison shows that the model that includes an interaction between length and stimulus. Thus length will be considered.


After finding out the best model that incorporates length is the one with an interaction with stimulus, I added model with a random slope for chiton. This model was not better than the model including an interaction with length but not a random slope. Therefore the model used was loomingmodel.example.


The best model is the one that includes an interaction between stimulus and length, but considering different response rates for each chiton does not make a difference. Thus loomingmodel.example is the best model.


I think from this I can conclude that the biggest effect of individual is on the positive control ("black"), but also has an impact on even background,compared with even black. Length slightly affects the responses of even black, but not much impact on +ive control or even background. 


Adding random slope for chiton (individual) did not make a difference, therefore this was excluded from the model and the above one is chosen.


```{r}
apred <- ggpredict(loomingmodel.example,'stimulus') 
error <- qt(0.975,df=n.count-1)*sd.count/sqrt(n.count)
xaxis <- c("Black", "even_background", "even_black", "White")

ggplot() + geom_point(data = apred, aes(x, predicted), colour = "green") + geom_errorbar(data = apred, aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), colour = "green") + geom_point(aes(apred$x, mean.count), colour = "black") + geom_errorbar(aes(xaxis, mean.count, ymin=mean.count-error, ymax=mean.count+error))
```

