---
title: "Threshold girdle"
author: "Chloe Brown"
date: "9/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library('sciplot')
library('multcomp')
library('ade4')
library('vegan')
library('ggplot2')
library('lme4')
library('tidyverse')
library('ggeffects')
```




```{r}
thresholdgirdle <- read_csv("thresholdgirdle.csv")
str(thresholdgirdle)
```




```{r}
mean.count <- tapply(thresholdgirdle$response,
                     list(thresholdgirdle$stimulus), 
                     mean)
mean.count
```


```{r}
plot(mean.count)
```


### The girdle model


```{r}
girdlemodel.4 <- glmer(
  response ~ stimulus + (1|stimulus_order) + (1|length) + (1|individual),
  data = thresholdgirdle, 
  family = binomial(), 
  control = glmerControl(optimizer = "optimx",
                         calc.derivs = FALSE,
                         optCtrl = list(
                           method = "nlminb", starttests = FALSE, kkt = FALSE)
                         )
  )

summary(girdlemodel.4)
```


### NUll model and model comparison

```{r}
girdlemodel.0000 <- glmer(
  response ~ 1 + (1|stimulus_order) + (1|length) + (1|individual),
  data = thresholdgirdle,
  family = binomial(),
  control = glmerControl(
    optimizer = "optimx", calc.derivs = FALSE,
    optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)
    )
  )

anova(girdlemodel.0000,girdlemodel.4)
```

### Make dummy data frame and plot predicted data

```{r}
newdata.df <- data.frame(
  # stimulus = with(thresholdgirdle, unique(stimulus)),
  stimulus   = rep(
                   seq(
                   from=min(thresholdgirdle$stimulus), 
                   to=max(thresholdgirdle$stimulus),
                   length.out = 200), # pick 200 pts of stimulus in the range
                   20),
  
  length     = rep(
                   seq(
                   from=min(thresholdgirdle$length), 
                   to=max(thresholdgirdle$length),
                   length.out = 20),
                   200),
  
  individual = sample(thresholdgirdle$individual, 4000,
                      replace = TRUE), # randomly pick ind
  
  stimulus_order = sample(thresholdgirdle$stimulus_order, 4000,
                      replace = TRUE)
  )

predictions <- predict(girdlemodel.4, #interval = 'confidence',
       newdata = newdata.df,
       allow.new.levels = TRUE
        )

plot(newdata.df$stimulus,plogis(predictions))
```

The above is primitive but might give you some idea of what is going on below. 


### Now get the marginal effects at the mean with the ggeffects package

Setting [n=500] or whatever number predicts many points in the range rather than just the original levels. 

```{r}
predz <- ggpredict(girdlemodel.4, 
                   terms = c('stimulus [n=500]') 
                   ) 
predz %>% plot()
```

### Get the turning point

Plot the derivative of the predicted line

```{r}
derivative = c(diff(predz$predicted) / diff(predz$x), 0)
find.df <- cbind.data.frame(predz$x,derivative)
ggplot(find.df, aes(predz$x, derivative)) + geom_line()
```

It's at the bottom as the line is descending. We get the value below.

```{r}
max_deriv <- which(derivative==min(derivative))
inflection_pt <- predz$x[max_deriv]
inflection_pt
```

Remake the plot with the point indicated.

```{r}
predz %>% plot() + geom_vline(xintercept = inflection_pt)
```



